6.830 Project Proposal \\ \large Data Imputation as a Query Optimization

### 2016-10-12 main.tex
proposal: added example table and diagrams

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}

@@

\section{Example}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage[margin=1in]{geometry}
\usepackage{qtree} % for diagram
\usepackage{listings} % for code excerpt
\usepackage{color} % for code excerpt
\lstset{
  basicstyle=\ttfamily\footnotesize,
  language=SQL,
  keywordstyle=\color{blue},
  otherkeywords={IMPUTE, impute}
}
\usepackage{booktabs} % for table

@@

\section{Example}
We consider a simple example that shows the potential impact of placing an imputation operator at different points in our query plan.

We assume the following relation \lstinline{Employees(id, salary, dept)}, with tuples shown in Table~\ref{example-table}. We impute by averaging over \lstinline{salary}, and we consider the query \lstinline{SELECT AVG(salary) FROM Employees GROUP BY dept}.

\begin{table}[h]
\centering
\caption{Example table with missing data}
\label{example-table}
\begin{tabular}{@{}lll@{}}
\toprule
id & salary & dept \\ \midrule
1  & 10     & 1    \\
2  & 15     & 1    \\
3  & NA     & 1    \\
4  & 5      & 2    \\ \bottomrule
\end{tabular}
\end{table}

\begin{figure}
 \begin{minipage}[b]{0.5\textwidth}
	\Tree [.\lstinline{AVG(salary)} [.\lstinline{GROUP BY (dept)} [.\lstinline{IMPUTE(Salary)} \lstinline{Employees} ] ] ]
    \caption{Early placement of impute operator}
    \label{early-fig}
  \end{minipage}
\hfill
 \begin{minipage}[b]{0.5\textwidth}
	\Tree [.\lstinline{AVG(salary)} [.\lstinline{IMPUTE(Salary)} [.\lstinline{GROUP BY (dept)} \lstinline{Employees} ] ] ]
    \caption{Late placement of impute operator}
    \label{late-fig}
  \end{minipage}
\end{figure}

In the early case (Figure~\ref{early-fig}), we impute over all salaries, and obtain as a final result \lstinline{[(1, 11.67), (2, 5)]}, while the late case (Figure~\ref{late-fig}) results in \lstinline{[(1, 12.5), (2, 5)]}.

This example shows that the imputation operator, even in its simplest form, can have a significant impact on the query results. Furthermore, this example fails to capture other important dimensions such as cost of imputation.
